Out-of-Sample Comparisons of Overfit Models

### 2014-03-28 paper.tex
Rough version of proof for in-sample statistics

\begin{thm}\label{res:insample1}
  Suppose Assumptions~\ref{asmp-1}--\ref{asmp-4} hold and let
  $\Lambda$ be a model selection statistic that takes the values zero
  or one; $\Lambda = 0$ indicates that the benchmark model is chosen
  and $\Lambda = 1$ indicates the alternative.  Moreover, assume that
  $L(e) = e^2$ and that there exist deterministic scalars $c$ and
  $\delta$ and a sequence of matrices $V_T$ such that
  \begin{equation}\label{eq:25}
    \limsup_{T \to \infty}
    \Pr[\Lambda \geq \1\{\bh{2T}'M_2' V_T M_2\bh{2T} > c \}]
    \leq \delta
  \end{equation}
  where the eigenvalues of $V_T$ are uniformly bounded, and $\lim
  \rank(V_T)/T > 0$.

@@

This behavior is formalized in Theorem~\ref{res:insample2}.
\begin{thm}\label{res:insample2}
  Suppose Assumptions~\ref{asmp-1}--\ref{asmp-4} hold, let $L(e) =
  e^2$, and let $\Lambda$ be a model selection statistic as in
  Theorem~\ref{res:insample1}. Also assume that, for any finite scalar
  $c$,
  \begin{equation}\label{eq:27}
    \Pr[\Lambda \leq \1\{\bh{2T}'M_2' V_T M_2\bh{2T} > c \}] \to 1
    \quad \text{as } T \to \infty,
  \end{equation}
  where $V_T$ is a sequence of deterministic matrices with uniformly
  bounded eigenvalues and $\lim \rank(V_T)/T > 0$. Then there exist
  DGPs satisfying these assumptions such that
  \begin{equation}\label{eq:15}
    \E( \Lambda \mid \E_T \oosB \geq 0 ) \to^p 0.
  \end{equation}
\end{thm}

@@

Let $(\e_{t+h}, x_{t}) \sim i.i.d.\ N(0,I)$ and let $\theta_1 = 0$.
For any $d \geq 0$, the event $\E_T \oosB = -d$ implies that
\begin{equation*}
  \E_T (y_{T+\h+1} - x_{1T+1}'\bh{1T})^2
  = \E_T (y_{T+\h+1} - x_{2T+1}'\bh{2T})^2 - d \quad a.s.
\end{equation*}
which can be expressed as
\begin{equation}\label{eq:34}
  d + \bh{1T}'\bh{1T} - \bh{2T}'M_1'M_1\bh{2T}
  = (\bh{2T} - \theta_2)' M_2'M_2 (\bh{2T} - \theta_2) \quad a.s.
\end{equation}
Now, $M_2 \bh{2T}$ is normally distributed conditional
on $\bh{1T}$ and $M_1 \bh{2T}$, and is distributed on the surface of
the sphere defined by~\eqref{eq:34} conditional on $\bh{1T}$, $M_1
\bh{2T}$, and the event $\E_T \oosB = -d$.

Since $M_2 \bh{2T}$ is normal, this conditional distribution is
invariant to reflection across the axes defined by the eigenvectors of
its covariance matrix. So when $\theta_2$ lies outside the cylindar
that contains the acceptance region of $\Lambda$, ie.e. $\theta_2'M_2'
V_T M_2 \theta_2 > c$, we have
\begin{align*}
  \Pr[ \bh{2T}'M_2' V_T M_2 &\bh{2T} \leq c \mid \E_T \oosB \leq 0] \\
  &= \E\Big( \Pr[ \bh{2T}'M_2' V_T M_2 \bh{2T} \leq c
  \mid \E_T \oosB \leq 0, \bh{1T}, M_1 \bh{2T}]
  \mid \E_T \oosB \leq 0 \Big) \\
  & < 1/2
\end{align*}
since the inner conditional probability is less than 1/2.

Then
\begin{equation*}
  \E(\Lambda \mid \E_T \oosB \leq 0)
  =   \E(\Lambda \mid \E_T \oosB \leq 0)
\end{equation*}

and magick...
\qed

@@

Then
\begin{equation*}
  \E( \Lambda \mid \E_T \oosB \geq 0 )
\end{equation*}

>>>>

\begin{thm}\label{res:insample1}
  Suppose Assumptions~\ref{asmp-1}--\ref{asmp-4} hold and let
  $\Lambda$ be a model selection statistic that takes the values zero
  or one; $\Lambda = 0$ indicates that the benchmark model is chosen
  and $\Lambda = 1$ indicates the alternative.  Moreover, assume that
  $L(e) = e^2$ and that there exist a deterministic scalar $c$
  and a sequence of possibly random matrices $V_T$ such that
  \begin{equation}\label{eq:25}
    \limsup_{T \to \infty}
    \Pr[\Lambda \geq \1\{\bh{2T}'M_2' V_T M_2\bh{2T} > c \}]
    \to 0
  \end{equation}
  where the eigenvalues of $V_T$ are uniformly bounded in probability,
  $\plim \rank(V_T)/T > 0$, and neither $V_T$ nor $c$ depend on
  $\theta$ or the value of $\bh{2T}$.

@@

This behavior is formalized in Theorem~\ref{res:insample2}.
\begin{thm}\label{res:insample2}
  Suppose Assumptions~\ref{asmp-1}--\ref{asmp-4} hold, let $L(e) =
  e^2$, and let $\Lambda$ be a model selection statistic as in
  Theorem~\ref{res:insample1}. Also assume that, for any finite scalar
  $c$,
  \begin{equation}\label{eq:27}
    \Pr[\Lambda \leq \1\{\bh{2T}'M_2' M_2\bh{2T} > c \}] \to 1
    \quad \text{as } T \to \infty.
  \end{equation}

  Then there exist DGPs satisfying these assumptions such that
  \begin{equation}\label{eq:15}
    \E( \Lambda \mid \E_T \oosB \geq 0 ) \to^p 0.
  \end{equation}
\end{thm}

@@

Let $(\e_{t+h}, x_{t}) \sim i.i.d.\ N(0,I)$ and let $\theta_1 = 0$.
For any $d \geq 0$, the event $\E_T \oosB = -d$ implies that
\begin{equation*}
  \E_T (y_{T+\h+1} - x_{1T+1}'\bh{1T})^2
  = \E_T (y_{T+\h+1} - x_{2T+1}'\bh{2T})^2 - d \quad a.s.
\end{equation*}
which can be expressed as
\begin{equation}\label{eq:34}
  d + \bh{1T}'\bh{1T} - \bh{2T}'M_1'M_1\bh{2T}
  = (\bh{2T} - \theta_2)' M_2'M_2 (\bh{2T} - \theta_2) \quad a.s.
\end{equation}
$M_2 \bh{2T}$ is normally distributed conditional on $\bh{1T}$ and
$M_1 \bh{2T}$, and is distributed on the surface of the sphere defined
by~\eqref{eq:34} conditional on $\bh{1T}$, $M_1 \bh{2T}$, and the
event $\E_T \oosB = -d$.

Since $M_2 \bh{2T}$ is normal, this conditional distribution is
invariant to reflection across the axes defined by the eigenvectors of
its covariance matrix. So when $\theta_2$ lies outside the cylindar
that contains the acceptance region of $\Lambda$, i.e. when
$\theta_2'M_2' V_T M_2 \theta_2 > c$, we have%
\footnote{To keep the notation in these equations managable, define
  the information set $\Gs = \sigma(\E_T \oosB \leq 0,~ \theta_2'M_2'
  V_T M_2 \theta_2 > c)$.} %
\begin{align*}
  \Pr[ \bh{2T}' M_2' V_T M_2 \bh{2T} \leq c \mid \Gs]
  &= \E\Big( \Pr[ \bh{2T}'M_2' V_T M_2 \bh{2T} \leq c \mid \Gs,~\bh{1T},~M_1 \bh{2T}] \mid \Gs \Big) \\
  & < 1/2
\end{align*}
since the inner conditional probability is less than 1/2.

Now let $\delta > 0$ be an arbitrary but small constant, and choose
$\theta_2$ far enough from the origin to ensure that
\begin{equation*}
  \Pr[\theta_2'M_2' V_T M_2 \theta_2 \leq c] \leq \delta
\end{equation*}
for large enough $T$. Then
\begin{align*}
  \E(\Lambda \mid \E_T \oosB \leq 0) &
  \geq \Pr[\bh{2T}' M_2' V_T M_2 \bh{2T} > c \mid \E_T \oosB \leq 0] \\
  &= \Pr[\bh{2T}' M_2' V_T M_2 \bh{2T} > c \text{~~and~~}
  \theta_2'M_2' V_T M_2 \theta_2 > c \mid \E_T \oosB  \leq 0] \\
  & \quad + \Pr[\bh{2T}' V_T M_2 \bh{2T} > c \text{~~and~~ }
  \theta_2'M_2' V_T M_2 \theta_2 \leq c \mid \E_T \oosB \leq 0].
\end{align*}

The second term is nonnegative by design. By conditioning on
$\theta_2'M_2' V_T M_2 \theta_2 > c$ our earlier argument shows that
the first term is greater than or equal to $(1 - \delta) / 2$. Since
$\delta$ is arbitrarily small, this completes the proof.
\qed

@@

Let $\delta$ be a small arbitrary constant, define $\theta^*$ as
\begin{gather*}
  \theta^* = \argmax_\theta \theta' M_2' M_2 \theta \\
  s.t.\ \bh{1T}'\bh{1T} - \bh{2T}'M_1'M_1\bh{2T} = (\theta - \theta_2)' M_2'M_2 (\theta - \theta_2)
\end{gather*}
($\theta^*$ depends on $\bh{1T}$ and $M_1 \bh{2T}$) and choose $c$ so
that
\begin{equation*}
  \Pr[ \theta^{* \prime} M_2 M_2 \theta^* > c] < \delta.
\end{equation*}

Then, we have for large enough $T$
\begin{align*}
  \E(\Lambda \mid \E_T \oosB \geq 0) &\leq
  \Pr[\bh{2T}'M_2'M_2\bh{2T} > c \mid \E_T \oosB \geq 0) \\
  &\leq \delta
\end{align*}
Since $\delta$ is arbitrarily close to zero, this completes the proof.
